---
title: "Juvenile Hare in Bonanza Creek Forest"
author: "Margaret Hammond"
date: "11/5/2021"
output:
  html_document:
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(broom)
library(equatiomatic)
library(lubridate)
library(kableExtra)
library(effsize)
```

# Data Introduction
In this document, the `bonanza_hares.csv` (Kielland et al. 2017) data is analyzed to investigate the relationship of hare weight and other variables. The data comes from the Bonanza Creek Experimental Forest in Fairbanks, Alaska. The data covers three separate collection sites focused on the different type of vegetation. These are placed into black spruce, mature, or riparian vegetation. The data was collected from 1999-2011, and focuses on the hare population. The data includes recordings on hare weight, hind foot length, age group, and sex. 


## Data and Analyses
In this report, the relationship between snowshoe hare trappings per year, snowshoe hare weight by gender, and snowshoe hare weight by hind feet length relationship are explored. The relationships are analyzed using summary tables, t-test's, Cohen's d, and Pearson's r. All relationship tests contain a segment analyzing their specific findings. Below is the data citation.

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

# Juvenile Hare Initial Investigation

## Annual Juvenile Hare Trap Counts
In this section, we will be analyzing the annual juvenile hare trap counts. The data will be rangled to include the total number of juvenile hare trappings during each year of the study. This data will then be compiled into a data visualization of the counts by year. 



```{r}
#Read in Data
#Read in the Bonanza Experimental Forest snowshoe hare data.
snowshoe <- read_csv(here("data", "bonanza_hares.csv"))
```

 

```{r}

# Wrangle the Data
# Wrangle the data to only have juvenile weight and create a year column.
juvenile_snowshoe <- snowshoe %>% 
  filter(age == "j") %>% # filter to only juvenile age
  mutate(date = mdy(date)) %>%  # adjust date column to date char
  mutate(year = year(date)) # create a year column

juvenile_snowshoe_count <- juvenile_snowshoe %>% 
  count(year)

snowshoe_trap_sum <- juvenile_snowshoe_count %>% 
  summarize( # Creating a summary table of weight by sex for the seperate sites
    mean_trap = mean(n),
    median_trap = median(n), # Median, sd, and n included for personal data observations
    sd_trap = sd(n), 
    min_trap = min(n),
    max_trap = max(n), 
    sample_size = n()
  )


```



```{r}

#Data Visualization
#Data visualization of juvenile snowshoe hare count by year. 
ggplot(data = juvenile_snowshoe_count, aes(x = year, y = n)) +
  geom_bar(stat = "identity",
           fill = "pink2", 
           color = "black") +
  labs(x = "Year",
       y = "Amount of Snowshoe Hares",
       title = "Juvenille Snowshoe Hare Count by Year",
       caption = "Figure 1. Juvenile Snowshoe hare count per year in Bonanza 
       Creek Experimental Forest from 1999 - present.")
```

### Data Visualization Observations
The data above shows the amount of juvenile snowshoe hare trappings per year in the Bonanza Creek Experimental Forest. The maximum amount trapped occurred in the year 1999, with `r snowshoe_trap_sum$max_trap[1]`. Of all the years that recorded trappings, the minimum amount trapped occurred in the year 2010 with `r snowshoe_trap_sum$min_trap[1]`. The mean amount of traps from 199-2012 were `r snowshoe_trap_sum$mean_trap[1]`, and the median amount was `r snowshoe_trap_sum$median_trap[1]`. As shown in `Figure 1` above, the notable trend of juvenile hare trap count is decreasing from 1999 to 2012. It is important to note that count is impacted by effort, and that earlier years could have had more amount of traps in the forest compared to later years. Moving forward, I suggest they standardized their trapping protocols to ensure the same amount of days spent trapping, and the same amount of traps. 


# Juvenile Hare Weight by Sex and Site

## Juvenile Hare Weight by Sex and Site Visualization 

```{r}
# Data Visualization
# Plot of juvenile hare weight by sex and site. Using the average weight of sex to explore the relationship.

weight_stats <- juvenile_snowshoe %>% 
  mutate(sex = case_when(
    sex == "f" ~ "F", # Change f to F for clearer organization
    sex == "m" ~ "M" # Change m to M for clearer organization
  )) %>% 
  mutate(grid = case_when(
    grid == "bonrip" ~ "Bonanza Riparian",
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonbs" ~ "Bonanza Black Spruce"
  )) 

weight_stats_sum <- weight_stats %>% 
  group_by(sex, grid) %>% 
  summarize( # Creating a summary table of weight by sex for the separate sites
    mean_weight = mean(weight, na.rm = TRUE),
    median_weight = median(weight, na.rm = TRUE), # Median, sd, and n included for personal data observations
    sd_weight = sd(weight, na.rm = TRUE), 
    sample_size = n()
  )


ggplot(data = weight_stats, 
       aes(x = sex, y = weight, fill = sex)) +
  geom_boxplot() +
  facet_wrap(~ grid, scale = "free") +
  scale_fill_manual(values = c("pink3", "blue", "grey")) + # pink for female, blue for male
 # theme()
  labs (x = "Sex",
        y = "Weight (gram)",
        title = "Weight (g) Juvenile Hare by Sex at each Site",
        caption = "Figure 2. Weight (g) of juvenile hare by 
        sex at the Bonanza Creek Experimental Forest from 1999-present.",
        fill = "Sex")
  
```

### Data Visualization Observations
`Figure 2` above shows the difference of weight between male, female, and `NA` sexes at the three Bonanza Creek collection sites; Black Spruce, Mature, and Riparian.  `NA` values are included for sex to allow for all data to be shown in the box plot. From `Figure 2`, we can see that the average male hare is heavier than the average female hare. This being said, different sites show large differences between average male and female hare weight, ~400 g. The Bonanza Mature site showed the largest difference between mean weight, whereas the Black Spruce site showed a smaller difference within ~200 g. 

# Juvenile Weight Comparison

```{r, include = FALSE}

#Exploratory Data Visualization
#Exploratory data visualization of the relationship between male and female juvenile hare weight. Done so with a QQ plot and histogram to explore assumptions of your t-test.
weight_compare_explore <- juvenile_snowshoe %>% 
  filter(sex != "NA") %>% 
  group_by(sex) %>% 
   mutate(sex = case_when(
    sex == "f" ~ "Female", #change f to Female for clearer organization
    sex == "m" ~ "Male")) # change m to Male for clearer organization

ggplot(data = weight_compare_explore, aes(x = weight)) +
  geom_histogram(bins = 15) + 
  facet_wrap(~sex)

ggplot(data = weight_compare_explore, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```


## Table Comparison of Male and Female weight


```{r}
#Compare mean weights for juvenile male and female snowshoe hares.

weight_compare <- juvenile_snowshoe %>% 
  filter(sex != "NA") %>% # Filter out NA for sex
  group_by(sex) %>% 
   mutate(sex = case_when(
    sex == "f" ~ "Female", # Change f to Female for cleaner table 
    sex == "m" ~ "Male")) %>% # Change m to Male for cleaner table
  summarize(
    mean_weight = 
      round(mean(weight, na.rm = TRUE), 2), # Use round() to show only 2 decimal places
    sd_weight = 
      round(sd(weight, na.rm = TRUE), 2),
    sample_size = n()
  )

weight_compare %>% 
  kable(col.names = c("Sex",
                      "Mean weight (g)", 
                      "Standard deviation (g)",
                      "Sample size (n)"),
        caption = "Figure 3. Juvenile Hare Weight (g) Comparison by Sex.") %>% 
  kable_styling(full_width = TRUE)

```


```{r}

# T-Test and Cohen's d
# Run a t-test and cohen's d to explore the relationship between female juvenile hare weight to male juvenile hare weight.
female_juvenile <- juvenile_snowshoe %>% 
  filter(sex == "f") %>% 
  drop_na(weight) %>%  
  pull(weight) # use pull to create a vector

male_juvenile <- juvenile_snowshoe %>% 
  filter(sex == "m") %>% 
  drop_na(weight) %>% 
  pull(weight) # use pull to create a vector


sex_weight_t <- t.test(x = female_juvenile, y = male_juvenile)

sex_weight_cohend <- cohen.d(female_juvenile, male_juvenile)

#sex_weight_t
#sex_weight_cohend
```


### Data Analysis
Mean weight for female hares (`r round(sex_weight_t$estimate[1],2)` mpg) is smaller than mean weight for male hares (`r round(sex_weight_t$estimate[2],2)` mpg). The difference of means between male and female hares is `r round(sex_weight_t$estimate[2],2) - round(sex_weight_t$estimate[1], 1)`. The effect size is small (Cohen's *d* = `r round(sex_weight_cohend$estimate,2)`) and the difference significant (two-sample t-test, t = `r round(sex_weight_t$statistic,2)`, p = .007).


# Juvenile Weight & Hind Foot Length


```{r}
##Wrangle Data
#Wrangle the data to include weight and hind foot length for juvenile hares.

weight_feet <- juvenile_snowshoe %>% 
  select(weight, hindft)
```

## Juvenile Weight by Hind Feet Visualization 

```{r}

#Data Visualization
#Data visualization of the relationship between weight and hind foot length in juvenile hares.
ggplot(data = weight_feet, 
       aes(x = hindft, y = weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Juvenile Hare Weight by Hind Feet Length",
       x = "Hind Foot Length (mm)",
       y = "Weight (g)",
       caption = "Figure 4. Juvenile hare hind feet (mm) by weight (g) in Bonanza Experimental Forest.") +
  theme_minimal()
```



```{r}

# Linear Regression
#Linear regression line to visualize how juvenile snowshoe hare weight changes with respect to hind foot length.

weight_feet_lm <- lm(hindft ~ weight, 
                 data = weight_feet)
weight_feet_lm_tidy <- tidy(weight_feet_lm) # tidy version of lm stats
weight_feet_lm_glance <- glance(weight_feet_lm) # even simplier version of lm stats

#weight_feet_lm_tidy
#weight_feet_lm_glance

weight_feet_cor <- cor.test(weight_feet$hindft, weight_feet$weight)
weight_feet_cor_tidy <- tidy(weight_feet_cor)
#weight_feet_cor_tidy

```

### Data Interpretation
On average, we expect a 1 mm increase in hind feet length to be associated with a `r round(weight_feet_lm_tidy$estimate[2], 2)` increase in body mass on average. The adjusted R^2^ value here (`r round(weight_feet_lm_glance$adj.r.squared, 2)`) indicates that `r 100*round(weight_feet_lm_glance$adj.r.squared, 2)`% of variance in body mass is explained by length of hind feet. Hind feet length and body mass are medium and positively correlated (Pearson's *r* = (`r round(weight_feet_cor_tidy$estimate, 2)`, p < 0.001). We are assuming normal distribution for weight and hind feet length distribution for juvenile hares. The assumption of homoscedasticity should be put into question, because inconsistent residuals exist within the model. Therefore, we might not want to use a linear regression model for data with inconsistent residuals.  


# Summary
In this data exploration, we have found:

 - Juvenile hare trappings have decreased from 1999-2011. 
 
 - Male juvenile hares weigh more, on average than female juvenile hares for all locations.
 
 - There is a moderate and positive correlation between hind feet length and juvenile hare weight.
 
# Citations
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

